RIScatter: Unifying Backscatter Communications, Symbiotic Radio, and Reconfigurable Intelligent Surface

### 2022-10-28 main.tex
:pencil: Update system model and achievable rates

	Simulation results demonstrate RIScatter nodes can exploit the additional propagation paths to smoothly transition between backscatter modulation and passive beamforming via smart input distribution design, and the proposed receiver significantly reduces the decoding complexity and effectively accommodates the double modulation and signal difference in scatter-based multiple access.

@@

		One special property of scatter-based multiple access is the backscatter signal strength is significantly weaker than primary due to the double fading effect.

@@

		\emph{Second,} we propose a practical receiver that accommodates the double modulation and signal difference in scatter-based multiple access.
		Since the primary and backscatter messages are superimposed by multiplication coding and the backscatter symbol is typically longer than primary, the scattered signals from RIScatter nodes can be treated as multipath components during primary decoding, and the primary message can be viewed as a spreading code during backscatter decoding.
		Conventional sequential primary-backscatter decoder eliminates primary interference by \gls{sic} at each primary block, while our sequential backscatter-primary detector semi-coherently decodes RIScatter nodes from the received energy, re-encodes to recover exact reflection patterns, and models the deterministic multipath within primary equivalent channel as dynamic passive beamforming at each backscatter block.
		It enables backscatter modulation and dynamic passive beamforming at much lower operational complexity, and is suitable for scatter nodes with various load switching speeds.

		\emph{Third,} we consider a scenario where multiple RIScatter nodes ride over an active point-to-point \gls{miso} transmission to perform backscatter modulation and passive beamforming towards a nearby user using shared spectrum, energy, and infrastructures.
		We provide primary and total backscatter rate analyses and characterize the achievable rate region by optimizing input distribution at RIScatter nodes, active beamforming at the \gls{ap}, and backscatter decision regions at the user.
		Since the original problem is highly non-convex, we decouple it into individual subproblems and propose a suboptimal \gls{bcd} algorithm, where the \gls{kkt} input distribution is numerically evaluated by limit of sequences, the active beamforming is iteratively updated by \gls{pgd} accelerated by \gls{bls}, and the decision regions are refined by existing sequential quantization methods for \gls{dmc}.
		This is the first paper to reveal the importance of input distribution and decision region designs in relevant literatures.

@@

\begin{section}{RIScatter}
	\begin{subsection}{Concepts}
		As a generalization of \gls{bc}, \gls{sr} and \gls{ris}, RIScatter is a passive scattering protocol that coexists with an active primary link in a flexible and mutualistic manner.
		RIScatter nodes utilizes \gls{csi}- and weight-based input distribution design to smoothly transition between backscatter modulation and passive beamforming.
		It can be implemented by adding an integrated receiver \cite{Kim2021a} and adaptive encoder \cite{He2020e} to off-the-shelf passive backscatter tags.
		The block diagram, equivalent circuit, and scatter model of a RIScatter node are illustrated in Fig. \ref{fg:riscatter_node}, while the input distribution and time structure comparison of scattering applications are shown in Fig. \ref{fg:scatter_comparison}.
		\begin{figure*}[!t]
			\centering
			\subfloat[Block Diagram]{
				\resizebox{0.32\linewidth}{!}{
					\input{assets/illustration/block_diagram.tex}
				}
				\label{fg:block_diagram}
			}
			\subfloat[Equivalent Circuit]{
				\resizebox{0.37\linewidth}{!}{
					\input{assets/illustration/equivalent_circuit.tex}
				}
				\label{fg:equivalent_circuit}
			}
			\subfloat[Scatter Model]{
				\resizebox{0.28\linewidth}{!}{
					\input{assets/illustration/scatter_model.tex}
				}
				\label{fg:scatter_model}
			}
			\caption{
			Block diagram, equivalent circuit, and scatter model of a RIScatter node.
			The solid and dashed vectors represent signal and energy flows.
			The scattering antenna behaves as a constant power source, where the voltage $V_0$ and current $I_0$ are introduced by incident electric field $\vec{E}_{\mathrm{I}}$ and magnetic field $\vec{H}_{\mathrm{I}}$ \cite{Huang2021}.
			}
			\label{fg:riscatter_node}
		\end{figure*}
		\begin{figure}[!t]
			\centering
			\subfloat[Input Distribution]{
				\resizebox{0.85\columnwidth}{!}{
					\input{assets/illustration/input_distribution.tex}
				}
				\label{fg:input_distribution}
			}
			\\
			\subfloat[Time Structure]{
				\resizebox{0.9\columnwidth}{!}{
					\input{assets/illustration/time_structure.tex}
				}
				\label{fg:time_structure}
			}
			\caption{
				Input distribution and time structure of \gls{bc}, \gls{ambc}, \gls{sr}, \gls{ris}, and RIScatter.
				``PB'' means primary block and ``BB'' means backscatter block.
			}
			\label{fg:scatter_comparison}
		\end{figure}
		Instead of using fully random or fully deterministic reflection pattern over time, each RIScatter node semi-randomly chooses the reflection state for each backscatter block with guidance of input probability distribution $P(\Gamma_m)$ at state $m$.
		Such an adaptive backscatter coding boils down to degenerate distribution of \gls{ris} when primary link is prioritized, and outperforms conventional line coding with equiprobable inputs (e.g., FM0 for \gls{rfid}) when backscatter link is prioritized.
		Besides, joint encoding over multiple co-located RIScatter nodes can further boost the total backscatter rate.
		Relevant \gls{csi} can be acquired by existing low-power estimation techniques.
		For dispersed RIScatter nodes, the cascaded transmitter-node-receiver channels can be estimated by sequential \cite{Bharadia2015,Yang2015b,Guo2019g} or parallel approaches \cite{Jin2021a} originally proposed for \gls{bc}.
		For co-located RIScatter nodes, the estimation can be simplified by group-based \cite{Zheng2019} and hierarchical \cite{You2019} trainings originally proposed for \gls{ris}.

		\begin{remark}
			Scatter-based multiple access involves a double modulation where the primary and backscatter symbols of different length are superimposed by multiplication coding.
			The reflection pattern not only encodes the backscatter message, but also affects the primary equivalent channel.
			Therefore, backscatter detection under primary uncertainty can be viewed as part of primary channel training, and novel decoding strategies apart from \gls{sic}
			\footnote{
				Recall that \gls{sic} was proposed to achieve the capacity vertices for \gls{mac} with superposition coding.
				For scatter-based multiple access, \gls{sic} not only fails to utilize the double modulation, but also requires $N$ re-encoding, precoding and subtraction with a time-domain \gls{mrc} during each backscatter block.
			}
			is desired for RIScatter.
		\end{remark}

		Next, we propose a cooperative receiver that effectively exploit signal characteristics (i.e., double modulation and symbol period difference) to reduce the decoding complexity and improve the primary-backscatter tradeoff.
		As illustrated in Fig. \ref{fg:energy_distribution}, conditioned on different reflection state hypotheses, the accumulated receive energy per backscatter block follows Gamma distribution with different scale parameters \cite{Qian2017b}.
		\begin{figure}[!t]
			\centering
			\resizebox{0.9\columnwidth}{!}{
				\input{assets/illustration/energy_distribution.tex}
			}
			\caption{
				\gls{pdf} of accumulated receive energy per backscatter block conditioned on different reflection state hypotheses.
				$z$, $t$, $\mathcal{H}$ and $\mathcal{R}$ denote the accumulated receive energy, decision threshold, reflection state hypothesis, and decision regions, respectively.
			}
			\label{fg:energy_distribution}
		\end{figure}
		Hence, the receiver can semi-coherently (in the presence of primary uncertainty) decode node messages from the accumulated receive energy, re-encode to recover actual reflection patterns, and model the deterministic multipath within primary equivalent channel.
		It enables backscatter modulation and dynamic passive beamforming using only one energy level comparison, re-encoding and precoding for each backscatter block (instead of primary block as \gls{sic}), and is suitable for arbitrary primary and backscatter symbol periods.
	\end{subsection}

	\begin{subsection}{System Model}
		\begin{figure}[!t]
			\centering
			\def\svgwidth{0.8\columnwidth}
			\footnotesize{
				\import{assets/illustration/}{riscatter_network.pdf_tex}
			}
			\caption{A single-user multi-node RIScatter system.}
			\label{fg:riscatter_network}
		\end{figure}
		To demonstrate the advantages of RIScatter, we consider a single-user multi-node \gls{miso} RIScatter system as shown in Fig. \ref{fg:riscatter_network}.
		In the primary active point-to-point system, a $Q$-antenna \gls{ap} transmits to a single-antenna user assisted by $K$ nearby single-antenna RIScatter nodes with $M$ reflection states.
		In the secondary backscatter \gls{mac} system, the \gls{ap} serves as the carrier emitter, the distributed RIScatter nodes modulate information over reradiated \gls{rf} signals, and the user decodes their messages.
		For simplicity, we consider a quasi-static block fading model where channels remain constant within coherence block and vary independently between consecutive blocks.
		Due to the physical constraints of load switching speed, we assume the backscatter over primary symbol period ratio is an integer $N \gg 1$.
		We also omit the signal reflected by two or more times and ignore the propagation time difference of different paths.

		Denote the \gls{ap}-user direct channel as $\boldsymbol{h}_{\mathrm{D}}^H \in \mathbb{C}^{1 \times Q}$, the \gls{ap}-node $k \in \mathcal{K} \triangleq \{1,\ldots,K\}$ forward channel as $\boldsymbol{h}_{\mathrm{F},k}^H \in \mathbb{C}^{1 \times Q}$, the node $k$-user backward channel as $h_{\mathrm{B},k}$, and the cascaded channel via tag $k$ as $\boldsymbol{h}_{\mathrm{C},k}^H \triangleq h_{\mathrm{B},k} \boldsymbol{h}_{\mathrm{F},k}^H \in \mathbb{C}^{1 \times Q}$.
		Let $x_{\mathcal{K}} \triangleq (x_1,\ldots,x_K)$ be the backscatter symbol tuple of all RIScatter nodes.
		Without loss of generality, we consider one specific backscatter block (i.e., $N$ primary blocks) in the following context.
		Due to double modulation, the primary equivalent channel is a function of \emph{coded} backscatter symbols\footnote{Expression \eqref{eq:equivalent_channel_bc} is often used in \gls{bc} literatures while expression \eqref{eq:equivalent_channel_ris} is commonly observed in \gls{ris} literatures.}
		% \begin{equation}
		% 	\boldsymbol{h}_{\mathrm{E}}^H(x_{\mathcal{K}}) \overset{\text{(a)}}{=} \boldsymbol{h}_{\mathrm{D}}^H + \sum_{k} \alpha_k \boldsymbol{h}_{\mathrm{C},k}^H x_k \overset{\text{(b)}}{=} \boldsymbol{h}_{\mathrm{D}}^H + \boldsymbol{x}^H \mathrm{diag}(\boldsymbol{\alpha}) \boldsymbol{H}_{\mathrm{C}},
		% 	\label{eq:equivalent_channel}
		% \end{equation}

@@

		where $\alpha_k \in \mathbb{I}$ is the amplitude scattering ratio of RIScatter node $k$, $x_k \in \mathcal{X} \triangleq \{c_1,\ldots,c_M\}$ is the \emph{coded} backscatter symbol of node $k$, $\boldsymbol{\alpha} \triangleq [\alpha_1,\ldots,\alpha_K]^T \in \mathbb{I}^{K}$, $\boldsymbol{x} \triangleq [x_1,\ldots,x_K]^H \in \mathcal{X}^{K}$, and $\boldsymbol{H}_{\mathrm{C}} \triangleq [\boldsymbol{h}_{\mathrm{C},1},\ldots,\boldsymbol{h}_{\mathrm{C},K}]^H \in \mathbb{C}^{K \times Q}$.

@@

		where $s \sim \mathcal{CN}(0,1)$ is the primary symbol, $v \sim \mathcal{CN}(0,\sigma_v^2)$ is the \gls{awgn}, and $\boldsymbol{w} \in \mathbb{C}^{Q}$ is the active beamforming vector subject to average transmit power constraint $\lVert \boldsymbol{w} \rVert^2 \le P$.
		Denote $m_k \in \mathcal{M} \triangleq \{1,\ldots,M\}$ as the reflection state index of node $k$, and $m_{\mathcal{K}} \triangleq (m_1,\ldots,m_K)$ as the state index tuple of all nodes.
		Conditioned on backscatter index tuple $m_{\mathcal{K}}$, the received signal at each primary block follows \gls{cscg} distribution $\mathcal{CN}(0,\sigma_{m_{\mathcal{K}}}^2)$, where

@@

		is the received variance and $x_{m_\mathcal{K}}$ is the backscatter symbol tuple associated with $m_{\mathcal{K}}$.
		For node $k$, let $x_{m_k}$ be the backscatter symbol associated with index $m_k$.%
		\footnote{
			Please note $x_k$ and $x_{\mathcal{K}}$ are random variable and tuple, while $x_{m_k}$ and $x_{m_{\mathcal{K}}}$ are their instances of index $m_k$ and $m_{\mathcal{K}}$.
		Besides, we define the total received energy per backscatter block as $z=\sum_{n \in \mathcal{N}} \bigl\lvert y[n] \bigr\rvert^2$.
		Since $z$ is the sum of $N$ \gls{iid} exponential variables, its \gls{pdf} conditioned on $m_{\mathcal{K}}$ follows Gamma distribution

@@

		As illustrated in Fig. \ref{fg:energy_distribution}, the RIScatter receiver divides the received energy space into disjoint decision regions associated with those hypotheses.
		For example, if the total received energy during a backscatter block falls within region $\mathcal{R}_{m_{\mathcal{K}}}$, then the detector output would be $x_{m_{\mathcal{K}}}$.

@@

			Interestingly, the capacity-achieving decision design for \gls{dmtc} remains under-investigated, and some attempts were made for a single source with binary inputs \cite{Qian2019b,Nguyen2021b}.
			For non-binary inputs with arbitrary distribution, the optimal decision region for each letter can be non-convex (i.e., with non-adjacent partitions) and the optimal number of thresholds is still unknown \cite{Nguyen2018,Nguyen2021}.

@@

		Like most existing literatures, we limit the scope of this paper to convex decision regions and consider sequential decision thresholds design therein.
		For the ease of notations, we define a general bijective function from backscatter index tuple $m_{\mathcal{K}}$ to integer $l \in \mathcal{L} \triangleq \{1,\ldots,L \triangleq M^K\}$, where $\{\sigma_l^2\}_{l \in \mathcal{L}}$ are sorted in an ascending order.
		Both notations are used interchangeably in the following context.
		As such, the convex decision region of backscatter tuple $l$ can be written as
			\mathcal{R}_{l} \triangleq [t_{l-1},t_l), \quad 0 \le t_{l-1} \le t_l.
		Once the decision threshold vector $\boldsymbol{t} \triangleq [t_0,\ldots,t_L]^T \in \mathbb{R}_+^{(L+1)}$ is determined, we can formulate a \gls{dmtmac} with transition probability from input $x_{m_{\mathcal{K}}}$ to output $\hat{x}_{m_{\mathcal{K}}'}$
			P(\hat{x}_{m_{\mathcal{K}}'}|x_{m_{\mathcal{K}}}) = \int_{\mathcal{R}_{m_{\mathcal{K}}'}} f(z|\mathcal{H}_{m_{\mathcal{K}}}) \, d z,
			\label{eq:dmtmac}
		then perform backscatter channel coding on top of it.
		\label{st:system_model}

	\begin{subsection}{Information Theory}
		Denote the input probability distribution vector of node $k$ as $\boldsymbol{p}_k \triangleq [P_k(c_1),\ldots,P_k(c_M)]^T \in \mathbb{I}^{M}$, and the probability of state $m_k$ as $P_k(x_{m_k})$.
		With independent encoding at all nodes, the probability of backscatter symbol tuple $x_{m_{\mathcal{K}}}$ is
		\begin{equation}
			P_{\mathcal{K}}(x_{m_{\mathcal{K}}}) = \prod_{k \in \mathcal{K}} P_k(x_{m_k}).
			\label{eq:equivalent_distribution}
		\end{equation}

>>>>

	Simulation results demonstrate RIScatter nodes can exploit the additional propagation paths to smoothly transition between backscatter modulation and passive beamforming via smart input distribution design, and the proposed receiver significantly reduces the decoding complexity and effectively accommodates the double modulation and signal difference in active-passive coexisting network.

@@

		One special property of active-passive coexisting network is the backscatter signal strength is significantly weaker than primary due to the double fading effect.

@@

		\emph{Second,} we propose a practical receiver that accommodates the double modulation and signal difference in active-passive coexisting network.
		Since the primary and backscatter messages are superimposed by multiplication coding and the backscatter symbol is typically longer than primary, the scattered signals from RIScatter nodes can be treated as multipath components during primary decoding, and the primary message can be viewed as a spreading code during backscatter decoding.
		Conventional sequential primary-backscatter decoder eliminates primary interference by \gls{sic} at each primary block, while our sequential backscatter-primary detector semi-coherently decodes RIScatter nodes from the received energy, re-encodes to recover exact reflection patterns, and models the deterministic multipath within primary equivalent channel as dynamic passive beamforming at each backscatter block.
		It enables backscatter modulation and dynamic passive beamforming at much lower operational complexity, and is suitable for scatter nodes with various load switching speeds.

		\emph{Third,} we consider a scenario where multiple RIScatter nodes ride over an active point-to-point \gls{miso} transmission to perform backscatter modulation and passive beamforming towards a nearby user using shared spectrum, energy, and infrastructures.
		We provide primary and total backscatter rate analyses and characterize the achievable rate region by optimizing input distribution at RIScatter nodes, active beamforming at the \gls{ap}, and backscatter decision regions at the user.
		Since the original problem is highly non-convex, we decouple it into individual subproblems and propose a suboptimal \gls{bcd} algorithm, where the \gls{kkt} input distribution is numerically evaluated by limit of sequences, the active beamforming is iteratively updated by \gls{pgd} accelerated by \gls{bls}, and the decision regions are refined by existing sequential quantization methods for \gls{dmtc}.
		This is the first paper to reveal the importance of input distribution and decision region designs in relevant literatures.

@@

\begin{section}{RIScatter}
	\begin{subsection}{Concepts}
		As a generalization of \gls{bc}, \gls{sr} and \gls{ris}, RIScatter is a passive scattering protocol that coexists with an active primary system in a flexible and mutualistic manner.
		RIScatter nodes leverage \gls{csi}- and weight-based input distribution design to smoothly transition between backscatter modulation and passive beamforming.
		It can be implemented by adding an integrated receiver \cite{Kim2021a} and adaptive encoder \cite{He2020e} to off-the-shelf passive backscatter tags.
		The block diagram, equivalent circuit, and scatter model of a RIScatter node are illustrated in Fig. \ref{fg:riscatter_node}, while the input distribution and time structure comparison of scattering applications are shown in Fig. \ref{fg:scatter_comparison}.
		\begin{figure*}[!t]
			\centering
			\subfloat[Block Diagram]{
				\resizebox{0.32\linewidth}{!}{
					\input{assets/illustration/block_diagram.tex}
				}
				\label{fg:block_diagram}
			}
			\subfloat[Equivalent Circuit]{
				\resizebox{0.37\linewidth}{!}{
					\input{assets/illustration/equivalent_circuit.tex}
				}
				\label{fg:equivalent_circuit}
			}
			\subfloat[Scatter Model]{
				\resizebox{0.28\linewidth}{!}{
					\input{assets/illustration/scatter_model.tex}
				}
				\label{fg:scatter_model}
			}
			\caption{
			Block diagram, equivalent circuit, and scatter model of a RIScatter node.
			The solid and dashed vectors represent signal and energy flows.
			The scattering antenna behaves as a constant power source, where the voltage $V_0$ and current $I_0$ are introduced by incident electric field $\vec{E}_{\mathrm{I}}$ and magnetic field $\vec{H}_{\mathrm{I}}$ \cite{Huang2021}.
			}
			\label{fg:riscatter_node}
		\end{figure*}
		\begin{figure}[!t]
			\centering
			\subfloat[Input Distribution]{
				\resizebox{0.85\columnwidth}{!}{
					\input{assets/illustration/input_distribution.tex}
				}
				\label{fg:input_distribution}
			}
			\\
			\subfloat[Time Structure]{
				\resizebox{0.9\columnwidth}{!}{
					\input{assets/illustration/time_structure.tex}
				}
				\label{fg:time_structure}
			}
			\caption{
				Input distribution and time structure of \gls{bc}, \gls{ambc}, \gls{sr}, \gls{ris}, and RIScatter.
				``PB'' means primary block and ``BB'' means backscatter block.
			}
			\label{fg:scatter_comparison}
		\end{figure}
		Instead of using fully random or fully deterministic reflection pattern over time, each RIScatter node semi-randomly chooses the reflection state for each backscatter block with guidance of input probability distribution $P(\Gamma_m)$ at state $m$.
		Such an adaptive backscatter coding boils down to degenerate distribution of \gls{ris} when primary link is prioritized, and outperforms conventional line coding with equiprobable inputs (e.g., FM0 for \gls{rfid}) when backscatter link is prioritized.
		Besides, joint encoding over multiple co-located RIScatter nodes can further boost the total backscatter rate.
		Relevant \gls{csi} can be acquired by existing low-power estimation techniques.
		For dispersed RIScatter nodes, the cascaded transmitter-node-receiver channels can be estimated by sequential \cite{Bharadia2015,Yang2015b,Guo2019g} or parallel approaches \cite{Jin2021a} originally proposed for \gls{bc}.
		For co-located RIScatter nodes, the estimation can be simplified by group-based \cite{Zheng2019} and hierarchical \cite{You2019} trainings originally proposed for \gls{ris}.

		\begin{remark}
			Scatter-based multiple access involves a double modulation where the primary and backscatter symbols of different length are superimposed by multiplication coding.
			The reflection pattern not only encodes the backscatter message, but also affects the primary equivalent channel.
			Therefore, backscatter detection under primary uncertainty can be viewed as part of primary channel training, and novel decoding strategies apart from \gls{sic}
			\footnote{
				Superposition coding and \gls{sic} was originally proposed to achieve the capacity vertices of Gaussian \gls{mac}.
				For active-passive coexisting network, \gls{sic} not only fails to utilize the multiplication coding/double modulation, but also requires $N$ re-encoding, precoding and subtraction with a time-domain \gls{mrc} during each backscatter block.
			}
			is desired for RIScatter.
		\end{remark}

		Next, we propose a cooperative receiver that effectively exploit signal characteristics (i.e., double modulation and symbol period difference) to reduce the decoding complexity and improve the primary-backscatter tradeoff.
		As illustrated in Fig. \ref{fg:energy_distribution}, conditioned on different reflection state hypotheses, the accumulated receive energy per backscatter block follows Gamma distribution with different scale parameters \cite{Qian2017b}.
		\begin{figure}[!t]
			\centering
			\resizebox{0.9\columnwidth}{!}{
				\input{assets/illustration/energy_distribution.tex}
			}
			\caption{
				\gls{pdf} of accumulated receive energy per backscatter block conditioned on different reflection state hypotheses.
				$z$, $t$, $\mathcal{H}$ and $\mathcal{R}$ denote the accumulated receive energy, decision threshold, reflection state hypothesis, and decision regions, respectively.
			}
			\label{fg:energy_distribution}
		\end{figure}
		Hence, the receiver can semi-coherently (in the presence of primary uncertainty) decode node messages from the accumulated receive energy, re-encode to recover actual reflection patterns, and model the deterministic multipath within primary equivalent channel.
		It enables simultaneous backscatter modulation and dynamic passive beamforming by only one energy level comparison, re-encoding and precoding for each backscatter block (instead of primary block), and is suitable for arbitrary primary sampling rate and backscatter switching speed.
	\end{subsection}

	\begin{subsection}{System Model}
			\def\svgwidth{0.8\columnwidth}
			\footnotesize{
				\import{assets/illustration/}{riscatter_network.pdf_tex}
			\caption{A single-user multi-node RIScatter system.}
			\label{fg:riscatter_network}
		As shown in Fig. \ref{fg:riscatter_network}, we consider an active-passive coexisting network where a $Q$-antenna \gls{ap} serves a single-antenna user and $K$ nearby dispersed and/or co-located single-antenna RIScatter nodes each with $M$ available states.
		In the primary point-to-point system, the \gls{ap} transmits information to the user over the multipath channel enhanced by RIScatter nodes.
		In the backscatter \gls{mac} system, the \gls{ap} and user become carrier emitter and backscatter reader, and the RIScatter nodes modulate over scattered \gls{rf} signals.
		For simplicity, we consider a quasi-static block fading model where channels remain constant within each block and vary independently between consecutive blocks, and assume the backscatter-over-primary symbol period ratio $N$ is an integer.
		We also omit the signal reflected by two or more times and ignore the propagation time difference of different paths.
		Denote the \gls{ap}-user direct channel as $\boldsymbol{h}_{\mathrm{D}}^H \in \mathbb{C}^{1 \times Q}$, the \gls{ap}-node $k \in \mathcal{K} \triangleq \{1,\ldots,K\}$ forward channel as $\boldsymbol{h}_{\mathrm{F},k}^H \in \mathbb{C}^{1 \times Q}$, the node $k$-user backward channel as $h_{\mathrm{B},k}$, and the cascaded \gls{ap}-node $k$-user channel as $\boldsymbol{h}_{\mathrm{C},k}^H \triangleq h_{\mathrm{B},k} \boldsymbol{h}_{\mathrm{F},k}^H \in \mathbb{C}^{1 \times Q}$.
		Let $x_k \in \mathcal{X} \triangleq \{c_1,\ldots,c_M\}$ be the coded backscatter symbol of node $k$ and $x_{\mathcal{K}} \triangleq (x_1,\ldots,x_K)$ be the backscatter symbol tuple of all nodes.
		Without loss of generality, we consider one backscatter block (i.e., $N$ primary blocks) in the following context.
		Due to double modulation, the primary equivalent channel is a function of backscatter symbol tuple\footnote{\eqref{eq:equivalent_channel_bc} and \eqref{eq:equivalent_channel_ris} are often used in \gls{bc} and \gls{ris} literatures, respectively.}

@@

		where $\alpha_k \in \mathbb{I}$ is the amplitude scattering ratio of node $k$, $\boldsymbol{\alpha} \triangleq [\alpha_1,\ldots,\alpha_K]^T \in \mathbb{I}^{K}$, $\boldsymbol{x} \triangleq [x_1,\ldots,x_K]^H \in \mathcal{X}^{K}$, and $\boldsymbol{H}_{\mathrm{C}} \triangleq [\boldsymbol{h}_{\mathrm{C},1},\ldots,\boldsymbol{h}_{\mathrm{C},K}]^H \in \mathbb{C}^{K \times Q}$.

@@

		where $\boldsymbol{w} \in \mathbb{C}^{Q}$ is the active beamformer satisfying average transmit power constraint $\lVert \boldsymbol{w} \rVert^2 \le P$, $s \sim \mathcal{CN}(0,1)$ is the primary symbol, and $v \sim \mathcal{CN}(0,\sigma_v^2)$ is the \gls{awgn} with average power $\sigma_v^2$.
		Let $m_k \in \mathcal{M} \triangleq \{1,\ldots,M\}$ be the reflection state index of node $k$, $m_{\mathcal{K}} \triangleq (m_1,\ldots,m_K)$ be the state index tuple of all nodes, $x_{m_k}$ be the backscatter symbol of node $k$ indexed by $m_k$, and $x_{m_\mathcal{K}}$ be the backscatter symbol tuple indexed by $m_{\mathcal{K}}$.\footnote{Please note $x_k$ and $x_{\mathcal{K}}$ are discrete random variable and tuple, while $x_{m_k}$ and $x_{m_{\mathcal{K}}}$ are their values indexed by $m_k$ and $m_{\mathcal{K}}$.}
		Conditioned on $m_{\mathcal{K}}$, the receive signal at each primary block follows \gls{cscg} distribution $\mathcal{CN}(0,\sigma_{m_{\mathcal{K}}}^2)$, where

@@

		is the received variance.
		Let $z=\sum_{n} \bigl\lvert y[n] \bigr\rvert^2$ be the accumulated receive energy per backscatter block.
		Since $z$ is the sum of $N$ \gls{iid} exponential variables, its conditional \gls{pdf} follows Gamma distribution

@@

		At the receiver, the energy space is divided into disjoint decision regions associated with different hypotheses, as illustrated in Fig. \ref{fg:energy_distribution}.

@@

			The capacity-achieving decision region design for \gls{dmtc} with non-binary inputs in arbitrary distribution remains an open issue.
			It was proved deterministic detector can be rate-optimal, but non-convex decision regions (i.e., comprise non-adjacent partitions) are generally required and the optimal number of thresholds is still unknown \cite{Nguyen2018,Nguyen2021}.
			Hence, we limit the backscatter energy detector to convex deterministic decision regions and consider sequential threshold design in the following context.

@@

		For the ease of notations, we map the state index tuple $m_{\mathcal{K}}$ to the corresponding index $l \in \mathcal{L} \triangleq \{1,\ldots,L \triangleq M^K\}$, where $\sigma_1^2,\ldots,\sigma_L^2$ is an ascending sequence.
		Both notations are used interchangeably in the following context.
		As such, the decision region of backscatter symbol tuple $l$ can be written as
		\begin{equation}
			\mathcal{R}_{l} \triangleq [t_{l-1},t_l), \quad 0 \le t_{l-1} \le t_l,
		\end{equation}
		where $t_l$ is the energy decision threshold between hypotheses $\mathcal{H}_{l}$ and $\mathcal{H}_{l+1}$.
		For a given decision threshold vector $\boldsymbol{t} \triangleq [t_0,\ldots,t_L]^T \in \mathbb{R}_+^{(L+1)}$, we can formulate a \gls{dmtmac} with transition probability from input $x_{m_{\mathcal{K}}}$ to output $\hat{x}_{m_{\mathcal{K}}'}$ given by
		\begin{equation}
			P(\hat{x}_{m_{\mathcal{K}}'}|x_{m_{\mathcal{K}}}) = \int_{\mathcal{R}_{m_{\mathcal{K}}'}} f(z|\mathcal{H}_{m_{\mathcal{K}}}) \, d z,
			\label{eq:dmtmac}
		\end{equation}
		over which adaptive input distribution design and backscatter channel coding can be performed.
		\label{st:system_model}
	\end{subsection}

	\begin{subsection}{Achievable Rates}
		Let $P_k(x_{m_k})$ be the input probability of node $k$ at state $m_k$, and $\boldsymbol{p}_k \triangleq [P_k(c_1),\ldots,P_k(c_M)]^T \in \mathbb{I}^{M}$ be the input distribution vector.
		% Denote the input probability of node $k$ at state $m_k$ as $P_k(x_{m_k})$, and the input distribution vector as $\boldsymbol{p}_k \triangleq [P_k(c_1),\ldots,P_k(c_M)]^T \in \mathbb{I}^{M}$.
		For dispersed nodes with independent encoding, the probability of backscatter symbol tuple $x_{m_{\mathcal{K}}}$ is
		\begin{equation}
			P_{\mathcal{K}}(x_{m_{\mathcal{K}}}) = \prod_{k \in \mathcal{K}} P_k(x_{m_k}).
			\label{eq:equivalent_distribution}
		\end{equation}
		Following \cite{Rezaeian2004}, we define the backscatter information function between input symbol tuple instance $x_{m_{\mathcal{K}}}$ and output symbol tuple variable $\hat{x}_{\mathcal{K}}$ as
		\begin{equation}
			I_{\mathrm{B}}(x_{m_{\mathcal{K}}};\hat{x}_{\mathcal{K}}) \triangleq \sum_{m_{\mathcal{K}}'} P(\hat{x}_{m_{\mathcal{K}}'}|x_{m_{\mathcal{K}}}) \log \frac{P(\hat{x}_{m_{\mathcal{K}}'}|x_{m_{\mathcal{K}}})}{P_{\mathcal{K}}(\hat{x}_{m_{\mathcal{K}}'})},
			\label{eq:backscatter_information_function}
		\end{equation}
		where $P_{\mathcal{K}}(\hat{x}_{m_{\mathcal{K}}'}) = \sum_{m_{\mathcal{K}}} P_{\mathcal{K}}(x_{m_{\mathcal{K}}}) P(\hat{x}_{m_{\mathcal{K}}'}|x_{m_{\mathcal{K}}})$ is the probability of output tuple $\hat{x}_{m_{\mathcal{K}}'}$.
		We also define the backscatter marginal information of letter $x_{m_k}$ as
		\begin{equation}
			I_{\mathrm{B},k}(x_{m_k};\hat{x}_{\mathcal{K}}) \triangleq \sum_{m_{\mathcal{K} \setminus \{k\}}} P_{\mathcal{K} \setminus \{k\}}(x_{m_{\mathcal{K} \setminus \{k\}}}) I_{\mathrm{B}}(x_{m_{\mathcal{K}}};\hat{x}_{\mathcal{K}}),
			\label{eq:backscatter_marginal_information}
		\end{equation}
		where $P_{\mathcal{K} \setminus \{k\}}(x_{m_{\mathcal{K} \setminus \{k\}}}) = \prod_{q \in \mathcal{K} \setminus \{k\}} P_{q}(x_{m_{q}})$.
		Hence, the backscatter mutual information can be written as
		\begin{equation}
			I_{\mathrm{B}}(x_{\mathcal{K}};\hat{x}_{\mathcal{K}}) = \sum_{m_{\mathcal{K}}} P_{\mathcal{K}}(x_{m_{\mathcal{K}}}) I_{\mathrm{B}}(x_{m_{\mathcal{K}}};\hat{x}_{\mathcal{K}}).
			\label{eq:backscatter_mutual_information}
		\end{equation}

		Once node messages are successfully decoded, we can re-encode for exact backscatter symbol tuple $x_{\mathcal{K}}$, recover their reflection patterns by \eqref{eq:backscatter_modulation}, and retrieve the primary equivalent channel by \eqref{eq:equivalent_channel}.
		Therefore, the primary information function conditioned on backscatter symbol tuple $x_{m_{\mathcal{K}}}$ is
		\begin{equation}
			I_{\mathrm{P}}(s;y|x_{m_{\mathcal{K}}}) \triangleq \log \Bigl(1 + \frac{\lvert \boldsymbol{h}_{\mathrm{E}}^H(x_{m_{\mathcal{K}}}) \boldsymbol{w} \rvert^2}{\sigma_v^2}\Bigr),
			\label{eq:primary_information_function}
		\end{equation}
		the primary marginal information of letter $x_{m_k}$ is
		% the primary marginal information conditioned on letter $x_{m_k}$ of node $k$ as
		\begin{equation}
			I_{\mathrm{P},k}(s;y|x_{m_k}) \triangleq \sum_{m_{\mathcal{K} \setminus \{k\}}} P_{\mathcal{K} \setminus \{k\}}(x_{m_{\mathcal{K} \setminus \{k\}}}) I_{\mathrm{P}}(s;y|x_{m_{\mathcal{K}}}),
			\label{eq:primary_marginal_information}
		\end{equation}
		and the primary ergodic mutual information is
		\begin{equation}
			I_{\mathrm{P}}(s;y|x_{\mathcal{K}}) = \sum_{m_{\mathcal{K}}} P_{\mathcal{K}}(x_{m_{\mathcal{K}}}) I_{\mathrm{P}}(s;y|x_{m_{\mathcal{K}}}).
			\label{eq:primary_mutual_information}
		\end{equation}
